---
title: "P8105 HW1"
author: "Veerapetch Petchger"
date: "2025-09-11"
output: github_document
---

## Problem 1
Load the moderndive library, and use the following code to load the early_january_weather dataset:
```{r setup, include = FALSE}
library(moderndive)
library(tidyverse)
data("early_january_weather")
```

```{r dataset}
early_january_weather
```

Write a short description of the dataset using inline R code; accessing the dataset help file can be informative. In your discussion, please include:

1) the variables in this dataset, including names / values of important variables

2) the size of the dataset (using nrow and ncol)

3) the mean temperature

This dataset has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns. It covers the hourly temperature measurements at Newark Liberty (`r (unique(early_january_weather$origin))`) International airport on from January `r min(early_january_weather$day)` to ` r max(early_january_weather$day)`. Important variables in this dataset include `temp` (temperature in Fahrenheit), `dewp` (dew point), `humid` (relative humidity), `wind_dir` (wind direction), `wind_speed`, `wind_gust`, `precip` (rainfall), `pressure`, and `visib` (visibility). The mean temperature over this period is `r mean(early_january_weather$temp)` °F.

Make a scatterplot of temp (y) vs time_hour (x); color points using the humid variable (adding color = ... inside of aes in your ggplot code should help). Describe patterns that are apparent in this plot.

```{r}

```


Export your scatterplot to your project directory using ggsave.



## Problem 2