---
title: "P8105 HW1"
author: "Veerapetch Petchger"
date: "2025-09-11"
output: github_document
---

## Problem 1
Load the moderndive library, and use the following code to load the `early_january_weather` dataset:
```{r setup, include = FALSE}
library(moderndive)
library(tidyverse)
data("early_january_weather")
```

```{r dataset}
early_january_weather
```

Write a short description of the dataset using inline R code; accessing the dataset help file can be informative. In your discussion, please include:

1) The variables in this dataset, including names / values of important variables

2) The size of the dataset (using nrow and ncol)

3) The mean temperature

This dataset has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns. It covers the hourly temperature measurements at Newark Liberty (`r (unique(early_january_weather$origin))`) International airport on from January `r min(early_january_weather$day)` to `r max(early_january_weather$day)`. Important variables in this dataset include `temp` (temperature), `dewp` (dew point), `humid` (relative humidity), `wind_dir` (wind direction), `wind_speed`, `wind_gust`, `precip` (rainfall), `pressure`, and `visib` (visibility). The mean temperature over this period is `r mean(early_january_weather$temp)` °F.

Make a scatterplot of temp (y) vs time_hour (x); color points using the humid variable (adding color = ... inside of aes in your ggplot code should help). Describe patterns that are apparent in this plot.

```{r}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point() +
  labs(
    title = "Early January Weather at Newark Liberty",
    subtitle = "Temperature vs Time, Points colored by relatie humidity",
    x = "Hour",
    y = "Air temperature",
    color = "Humidity",
  )
```
The temperatures generally rise during the daytime and decline overnight. Humidity appears to have an inverse relationship with weather, recording its highest values during the night. Over the span of days, as the daily temperatures slightly increase, humidity also tends to be higher.

Export your scatterplot to your project directory using ggsave.

## Problem 2